%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Master Thesis - Computer Engineering
%% Copyright 2009 Ricardo Alexandre Fiorelli, Erick Poletto
%% This document is distributed by the terms of the license
%% included in the file LICENCE.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Acknowledgements
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\chapter*{Acknowledgements}
% Agradecimentos - é so para as pessoas que contribuiram relevantemente
% para a elaboração do trabalho
To professor Chiara Francalanci for her advice and guidance in this research as well as the extraordinary experiences throughout the work.

To professor Paolo Giacomazzo, for the great ideas and enthusiastic comments. Also, for the well given critics in the early stage of the research project.

To Francesco Merlo, for the great help in the final stages of the project and for the guidance for the production of this document.

To our friends Pedro Kayatt, Jobson Wadi, Rafael Sbampato, Rodrigo Roriz, Xenia Itikawa, Sarp Erdag, Lais Paixao, Guilherme Matiussi, Rafael Vilella, Lucas de Marchi, Emilio Rosseto, Rafaella Martins, Helio Omoto and Cristina Chang who provided their time in order to make possible the field benchmarking and power measurements.

\subparagraph*{Erick}
To my family, who have always provided support and guidance. To my parents Annie and Marcos, who have taught me to act with tranquility and balance. To my brothers Henry, William and my sister Mariana, for always cheering me up. And to my friends, whom I hold very dear and which have always been by my side.

\subparagraph*{Ricardo}
To my father, for always having something to support me. To my mother and sister Priscilla, for taking a long trip from Brazil just to participate in my discussion. To my other sister Lilian, who wishes she could have come. To my girlfriend, for being an amazing person always. And to my friends who helped and supported me in some way throughout the work.
