%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Master Thesis - Computer Engineering
%% Copyright 2009 Ricardo Alexandre Fiorelli, Erick Poletto
%% This document is distributed by the terms of the license
%% included in the file LICENCE.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Appendix C
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\chapter{List of Other Energy Management Tools} \label{app:list_other_energy_management_tools}
    \section{Power To Change} \label{app:power_to_change}
    Power To Change is a widget for desktops that measures how much energy was saved when the computer is turned-off. With this application installed, when the machine is turned on, the user can receive information about how much energy and carbon footprint it was saved while it was turned off, and also, compare with global results and others.
    The widget can be downloaded from \url{http://www.hp.com/powertochange}.

    \section{PlateSpin - Recon} \label{app:power_recon}
    This software did not compose the ones used for doing this thesis. Yet, it is important to notice this, because it is almost the same of Sandra, but it provides a more incisive work on Data Centers in general. It provides workload profiling, analysis and planning of complex server consolidation, disaster recovery, capacity planning, asset management and green data center initiatives. It also provides forecasting for optimizing the data center by collecting hardware, software and services inventory for all server workloads. Furthermore, it results an statistics work for the server workloads running on data center and how their resources are being used.
    
    \section{APC Virtualization Energy Cost Calculation}\label{app:apc_virtualization_energy_cost_calculation}
        APC has released a free online tool that will enable managers make decisions regarding the physical requirements of a data center after a virtualization project. The Virtualization Energy Cost Calculator, which can be downloaded from the APC site, can be used in data center planning and allows the simulation of different scenarios.

        It does not, however, allow to differentiate between different servers and their power consumption. ``It is not meant as an accurate planning tool'', said an APC spokesman. ``The idea is that gives facilities and IT managers a rough idea of the power consumption but it does not go into detail. APC does have another product to get a more accurate picture - InfrastuXure Central, although that is not a free tool''.
        %\url{http://www.techworld.com/green-it/news/index.cfm?RSS&NewsID=116650}
        
        
